Variance reduction for MCMC methods via martingale representations

### 2020-01-19 aap/Revision/vr-mcmc-revision.tex
update

%\som{Changed exponent from 1/2 to 2 here}
Now, from Hoelder inequality, we obtain (with $\|X\|_p = (\mathsf{E}X^p)^{\frac{1}{p}}$)
\begin{align*}
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left\Vert \sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}^{2}\\
 & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\left\Vert \prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}\right)^{2}.
\end{align*}
Now using the fact that $\prod\limits_{k=j+1}^{p}\alpha_{k} \leq \exp\left(-\sum\limits_{k=j+1}^{p}\gamma_{k}\mu'(X_{k-1})\right),$
we get 
\[
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right]\leq\sum_{j=l}^{q}\gamma_{j} \zeta^{1/2}_{l,j}(2) \left(\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4)\right)^{2},
\]
where we denote
\[
\zeta_{l,j}(s)=\mathsf{E}\left[e^{-s\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})}\right],\quad s>0.
\]
From Theorem~\ref{exp_mom_ula} in Appendix~\ref{sec:appendix}, it follows that
\[
\zeta_{l,j}(s) \leq e^{s\kappa \Gamma^*}e^{-2s\Gamma_{l,j}\alpha}
\]
with $\Gamma^* = \sum\limits_{k=1}^{\infty}\gamma_k^2, \kappa = ..., \Gamma_{l,j} = \sum\limits_{k=l}^{j}\gamma_k, \alpha = \int \mu^2(x)\pi(x)\,dx$. Since $\sum\limits_{k=1}^{\infty}\gamma_k = \infty$, we may bound
\begin{align*}
\sum\limits_{p=j+1}^{\infty} \gamma_p e^{-2\alpha \Gamma_{j+1,p}} < \int\limits_{0}^{+\infty}e^{-2\alpha x}\,dx = \frac{1}{2\alpha}
\end{align*}
yielding 
\begin{align*}
\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4) \leq e^{\kappa \Gamma^*}\sum_{p=j+1}^{q+1}\gamma_{p}e^{-2\alpha \Gamma_{j+1,p}} \leq \frac{e^{\kappa \Gamma^*}}{2\alpha} 
\end{align*}
Using the same integral bound, 
\begin{align*}
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] \leq
\frac{e^{2\kappa \Gamma^*}}{2\alpha}\sum_{j=l}^{q}\gamma_{j}e^{-2\alpha \Gamma_{l,j}} \leq \frac{e^{2\kappa \Gamma^*}}{4\alpha^2}
\end{align*}
Plugging into bound for conditional variance, 
\begin{align*}
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right] \leq
\frac{A_{K}^{2}e^{2\kappa \Gamma^*}}{4\alpha^2(1-\gamma_1 B_\mu)^3}
\end{align*}
\par
The proof is completed.
\end{proof}
\begin{remark}\label{rem:p-norm}
In fact in the proof of Lemma~\ref{lem:var_poincare} using \eqref{eq:part-part} we have shown that the expectations
\begin{eqnarray*}
\mathsf E\left[\left.
\|\nabla_Z F(X_{l-1},Z_l,Z_{l+1},\ldots,Z_q)\|^r
\right|X_{l-1}\right]
\end{eqnarray*}
are uniformly bounded for all natural \(l\leq q\) and any \(r\geq 2.\) Indeed it follows from \eqref{eq:17042018a1} that  the quantities
\begin{eqnarray*}
\left(\sum_{j=l}^{q} \gamma_{j}
\prod_{k=l}^{j} \alpha_{k}\right)^{r/2}
\left(
\sum_{p=j+1}^{q+1} \gamma_{p} \prod_{k=j+1}^{p} \alpha_{k}
\right)^{r}
\end{eqnarray*}
are bounded. This stronger result in combination with a \(L^r\)-Sobolev-type inequalities (see e.g. \cite{Adamczak}) implies that also higher order conditional (on \(X_{l-1}\)) central moments of the r.v.
\begin{eqnarray*}
\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)
\end{eqnarray*}
are uniformly bounded in \(l\leq q\) and \(p.\)

>>>>

Now, from Hoelder inequality, we obtain (with $\|X\|_p = (\mathsf{E}X^p)^{\frac{1}{p}}$)
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left\Vert \sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}^{2}\\
 & \leq\sum_{j=l}^{q}\gamma_{j}\left\Vert \prod_{k=l}^{j}\alpha_{k}\right\Vert _{2}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\left\Vert \prod_{k=j+1}^{p}\alpha_{k}\right\Vert _{4}\right)^{2}.
Now using the fact that $\prod\limits_{k=j+1}^{p}\alpha_{k} \leq \exp\left(-\sum\limits_{k=j+1}^{p}\gamma_{k}\mu'(X_{k-1})\right),$
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right]\leq\sum_{j=l}^{q}\gamma_{j} \zeta^{1/2}_{l,j}(2) \left(\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4)\right)^{2},
where we denote
\zeta_{l,j}(s)=\mathsf{E}\left[e^{-s\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})}\right],\quad s>0.
From Theorem~\ref{exp_mom_ula} in Appendix~\ref{sec:appendix}, it follows that
\mathsf{E}\left[e^{-s\left(\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})-\Gamma_{l,j}\int\mu'(x)\pi(x)\,dx\right)}\right]\leq e^{s\kappa\sum_{k=l}^{j}\gamma_{k}^{2}}
with constant $\kappa = ...$. We used the fact that $\mu^{\prime}(x)$ is bounded. Denote $\sum_{k=1}^{\infty}\gamma_{k}^{2} = \Gamma^{\star}$, then we have 
\mathsf{E}\left[e^{-s\sum_{k=l}^{j}\gamma_{k}\mu'(X_{k-1})}\right]\leq e^{s\varkappa\Gamma^{\star}}e^{-s\Gamma_{l,j}\int\mu'(x)\pi(x)\,dx}.
Furthermore, since $\mu\pi',\mu'\pi\in L^{1}(\mathbb{R})$, we have
\int\mu'(x)\pi(x)\,dx=-\int\mu(x)\pi'(x)\,dx=2\int\mu^{2}(x)\pi(x)\,dx>0
yielding the final bound
\zeta_{l,j}(s) \leq e^{s\kappa \Gamma^*}e^{-2s\Gamma_{l,j}\alpha}
where $\Gamma_{l,j} = \sum\limits_{k=l}^{j}\gamma_k, \alpha = \int \mu^2(x)\pi(x)\,dx$. Since $\sum\limits_{k=1}^{\infty}\gamma_k = \infty$, we may estimate
\sum\limits_{p=j+1}^{\infty} \gamma_p e^{-2\alpha \Gamma_{j+1,p}} < \int\limits_{0}^{+\infty}e^{-2\alpha x}\,dx = \frac{1}{2\alpha}
yielding 
\sum_{p=j+1}^{q+1}\gamma_{p} \zeta^{1/4}_{j+1,p}(4) \leq e^{\kappa \Gamma^*}\sum_{p=j+1}^{q+1}\gamma_{p}e^{-2\alpha \Gamma_{j+1,p}} \leq \frac{e^{\kappa \Gamma^*}}{2\alpha} 
Using the same integral bound, 
\mathsf{E}\left[\sum_{j=l}^{q}\gamma_{j}\prod_{k=l}^{j}\alpha_{k}\left(\sum_{p=j+1}^{q+1}\gamma_{p}\prod_{k=j+1}^{p}\alpha_{k}\right)^{2}\right] \leq
\frac{e^{2\kappa \Gamma^*}}{2\alpha}\sum_{j=l}^{q}\gamma_{j}e^{-2\alpha \Gamma_{l,j}} \leq \frac{e^{2\kappa \Gamma^*}}{4\alpha^2}
Plugging into bound for conditional variance, 
\mathsf{Var}\left[\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)\Big|X_{l-1}\right] \leq
\frac{A_{K}^{2}e^{2\kappa \Gamma^*}}{4\alpha^2(1-\gamma_1 B_\mu)^3}
The proof is completed.
\begin{remark}\label{rem:p-norm}
In fact in the proof of Lemma~\ref{lem:var_poincare} using \eqref{eq:part-part} we have shown that the expectations
\mathsf E\left[\left.
\|\nabla_Z F(X_{l-1},Z_l,Z_{l+1},\ldots,Z_q)\|^r
\right|X_{l-1}\right]
are uniformly bounded for all natural \(l\leq q\) and any \(r\geq 2.\) Indeed it follows from \eqref{eq:17042018a1} that  the quantities
\left(\sum_{j=l}^{q} \gamma_{j}
\prod_{k=l}^{j} \alpha_{k}\right)^{r/2}
\left(
\sum_{p=j+1}^{q+1} \gamma_{p} \prod_{k=j+1}^{p} \alpha_{k}
\right)^{r}
are bounded. This stronger result in combination with a \(L^r\)-Sobolev-type inequalities (see e.g. \cite{Adamczak}) implies that also higher order conditional (on \(X_{l-1}\)) central moments of the r.v.
\sum_{p=l}^{q}\gamma_{p+1}\partial_{y_{l}}^{K}f\left(X_{p}\right)
are uniformly bounded in \(l\leq q\) and \(p.\)
